---
title: "Grace Li_A3"
author: Grace Li
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r}
library(censusapi)
library(tidyverse)
library(tigris)
library(sf)
library(leaflet)

Sys.setenv(CENSUS_KEY="9136d9427440997e3a0e797bbf837fa633066a90")
```

```{r}
temp <- tempfile()
download.file("https://www2.census.gov/programs-surveys/acs/data/pums/2019/1-Year/csv_hca.zip",destfile = temp)

pums_hca_2019_1yr <- read_csv(unzip(temp,"psam_h06.csv"))

unlink(temp)
```


```{r}
pums_2019_1yr <- getCensus(
  name = "acs/acs1/pums",
  vintage = 2019,
  region = "public use microdata area:*", 
  regionin = "state:06",
  vars = c(
    "SERIALNO",
    "SPORDER",
    "PWGTP",
    "WGTP",
    "NP",
    "HHL",
    "HINCP",
    "AGEP",
    "R65"
  )
)
```


```{r}
ca_pumas <-
  pumas("CA", cb = T, progress_bar = F)

bay_county_names <-
  c(
    "Alameda",
    "Contra Costa",
    "Marin",
    "Napa",
    "San Francisco",
    "San Mateo",
    "Santa Clara",
    "Solano",
    "Sonoma"
  )

bay_counties <-
  counties("CA", cb = T, progress_bar = F) %>%
  filter(NAME %in% bay_county_names)

bay_pumas <-
  ca_pumas %>% 
  st_centroid() %>% 
  .[bay_counties, ] %>% 
  st_drop_geometry() %>% 
  left_join(ca_pumas %>% select(GEOID10)) %>% 
  st_as_sf()

bay_pums <-
  pums_2019_1yr %>% 
  mutate(
    PUMA = str_pad(public_use_microdata_area,5,"left","0")
  ) %>% 
  filter(PUMA %in% bay_pumas$PUMACE10)
```

```{r}
bay_pums_elder <-
  bay_pums %>% 
  mutate(
    WGTP = as.numeric(WGTP),
    nonenglish_elder_lowinc = ifelse(
      (HHL != 1) &
        (AGEP > 60) &
        (HINCP < 100000),
      WGTP,
      0
    )
  ) %>% 
  group_by(PUMA) %>% 
  summarize(
    perc_nonenglish_elder_lowinc =
      sum(nonenglish_elder_lowinc, na.rm =T)/sum(WGTP, na.rm = T)*100
  ) %>% 
  left_join(
    bay_pumas %>% 
      select(PUMACE10),
    by = c("PUMA" = "PUMACE10")
  ) %>% 
  st_as_sf()
```


```{r}
pums_pal <- colorNumeric(
  palette = "Oranges",
  domain = bay_pums_elder$perc_nonenglish_elder_lowinc
)

leaflet() %>%
  addTiles() %>% 
  addPolygons(
    data = bay_pums_elder,
    fillColor = ~pums_pal(perc_nonenglish_elder_lowinc),
    color = "white",
    opacity = 0.5,
    fillOpacity = 0.5,
    weight = 1,
    label = ~paste0(
      round(perc_nonenglish_elder_lowinc), 
      "% non-English-speaking Elderly(Over 60) Making Less Than $100K"
    ),
    highlightOptions = highlightOptions(
      weight = 2,
      opacity = 1
    )
  ) %>% 
  addLegend(
    data = bay_pums_elder,
    pal = pums_pal,
    values = ~perc_nonenglish_elder_lowinc,
    title = "% non-English-speaking<br>Elderly (Over 60)<br>Making Less Than $100K"
  )
```

# Reflection

From my interview designed on last Tuesday, we discussed how older people might be more vulnerable to heat waves. They are more likely to be affected because of their health condition. Meanwhile, non-English-speaking low income people are those more disadvantaged in our community, so they might lack the resources to deal with such natural disaster. This is why I chose "non-English-speaking elderly with income less than $100K" as a vulnerable group for discussion.

From the interactive graph, we can see that percentages vary across counties, ranging from 1% to the most severe 10%. I believe local authorities should pay more attention to the blocks that generally have a darker color, since those places will have more elderly low income people adversely affected by heat waves.

