---
title: "Grace_Li_A6_new"
author: "Grace Li"
date: "12/13/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r library}
library(censusapi)
library(tidyverse)
library(tigris)
library(sf)
library(leaflet)
Sys.setenv(CENSUS_KEY="9136d9427440997e3a0e797bbf837fa633066a90")
```

```{r metadata,eval=F}
pums_2019_1yr <- getCensus(
  name = "acs/acs1/pums",
  vintage = 2019,
  region = "public use microdata area:*", 
  regionin = "state:06",
  vars = c(
    "SERIALNO",
    "SPORDER",
    "PWGTP",
    "WGTP",
    "YBL",
    "BLD",
    "TEN",
    "MV",
    "NP",
    "HINCP",
    "AGEP"
  )
)
saveRDS(pums_2019_1yr, "a6_pums.rds") 
```

```{r metadata-loading}
pums_2019_1yr <-readRDS("a6_pums.rds")
```

```{r motify-data-for-sf}
ca_pumas <-
  pumas("CA", cb = T, progress_bar = F)
sf_boundary <-
  counties("CA", cb = T, progress_bar = F) %>%
  filter(NAME == "San Francisco")
sf_pumas <-
  ca_pumas %>% 
  st_centroid() %>% 
  .[sf_boundary, ] %>% 
  st_drop_geometry() %>%
  left_join(ca_pumas %>% select(GEOID10)) %>% 
  st_as_sf()
sf_pums <-
  pums_2019_1yr %>% 
  mutate(
    PUMA = str_pad(public_use_microdata_area,5,"left","0")
  ) %>% 
  filter(PUMA %in% sf_pumas$PUMACE10)
``` 

```{r map-sf-pumas-1}
leaflet() %>% 
  addTiles() %>% 
  addPolygons(
    data = sf_pumas
  ) %>%
  addMarkers(
    data = sf_pumas %>% 
      st_centroid()
  )
```

```{r map-sf-pumas-2, eval =F}
ggplot(sf_pumas) + geom_sf()
```

```{r}
sf_pums_clean <- sf_pums %>%
  filter(YBL %in% c("1","2","3")) %>% 
  mutate(
    YBL = as.numeric(YBL),
    HINCP = as.numeric(HINCP),
    AGEP = as.numeric(AGEP)
         ) %>%
  group_by(SERIALNO, .drop = FALSE) %>%
  arrange(AGEP) %>% 
  summarise_all(first) %>%
  mutate(
    LR = case_when(
          AGEP < 6 & HINCP < 90000~1,
          TRUE ~ 0
    )
  )
```

```{r factoring}
sf_pums_clean <- sf_pums_clean %>% 
  mutate(
    BLD = as.factor(sf_pums_clean$BLD),
    TEN = as.factor(sf_pums_clean$TEN),
    MV = as.factor(sf_pums_clean$MV),
    PUMA = as.factor(sf_pums_clean$PUMA)
  )
```


```{r}
pums_model <- glm(LR ~BLD+TEN+MV+PUMA, family = quasibinomial(), sf_pums_clean)
```

```{r}
summary(pums_model)
```

```{r cof-table}
cof2 <- summary(pums_model)$coefficients
knitr::kable(cof2,
             caption = "Cofficient Table",
             digits = 2)
```

```{r predict}
new_data = sample_n(sf_pums_clean, size = 1)
```

The predicted score of risk with random selection of data is `r predict(pums_model, new_data, type ="response")` , where its true value is `r new_data$LR` .


```{r}
sf_pums_clean <- sf_pums_clean %>%
  cbind(
    SCORE = predict(pums_model,sf_pums_clean, type = "response")
  )
```

```{r}
summary_2x2 <-
  sf_pums_clean %>% 
  mutate(
    LR = ifelse(
      LR == 1, 
      "Yes (Risk)", 
      "No (No Risk)"
    )
  ) %>% 
  pull(LR) %>% 
  table(sf_pums_clean$SCORE >= 0.1)
```

```{r error-table}
knitr::kable(summary_2x2, caption = "Type1/2 Table")
```

The households that exceeded my threshold score are generally one family houses occupied for 5-9 years without payment of rent and built in 1939 or earlier. 

The success rate is relatively low. Roughly 31 sets of postcards (not considering WGTP) will be mailed out based on the specified threshold. Only 13% of the postcards actually go to low-income households with children, and 11% of children at risk are likely to be reached. Possible improvement to the strategy can be lowering the threshold value to improve type1 error, even though type2 error will also be affected.


```{r}
temp <- tempfile()
download.file("https://www2.census.gov/programs-surveys/acs/data/pums/2019/1-Year/csv_hca.zip",destfile = temp)
pums_hca_2019_1yr <- read_csv(unzip(temp,"psam_h06.csv"))
unlink(temp)
```


```{r}
test <- pums_hca_2019_1yr %>%
  group_by(SERIALNO) %>%
  select(starts_with("WGTP"))
sf_pums_clean <- sf_pums_clean %>%
  left_join(., test, by = "SERIALNO")
```

```{r}
model2 <- glm(LR ~BLD+TEN+MV+PUMA, family = quasibinomial(), sf_pums_clean)
```

```{r}
summary_rw <-
  sf_pums_clean %>% 
  select(LR, SCORE, starts_with("WGTP"), -WGTP.x, -WGTP.y) %>%
  group_by(LR, SC=SCORE>0.1) %>%
  summarise_all(sum)
```



